Using WordNet and Short-Term Memory for Contextual Disambiguation

### 2017-02-24 Draft_1.tex
Added more detail to description implementation section. Updated to reflect recent changes to disambiguation, i.e. delay and hypernym model

\documentclass[]{article}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{tcolorbox}

@@

\begin{itemize}
	\item \textbf{Document}, is a list of paragraphs,
	\item \textbf{Paragraphs}, each of which is a list of sentences,
	\item \textbf{Sentences}, each of which is a list of words.
\end{itemize}

@@

@@

Both an STS and Episodic Buffer have been implemented, using python classes for each. Interactions between the two memory structures are handled by a Memory Controller.

@@

The purpose of the episodic buffer, is to maintain a list of synsets which have occurred in the STS previously. This allows future activations of synsets which have already occured to receive a boost when activated in the future. Interactions with the Episodic buffer only occcur during the activation of synsets, therefore, all interactions are handled by the Meory Controller.

@@

The corpus is read a sentence at a time, for each word in the sentence a series of activations takes place, according to the pseudo-code shown in Listing \ref{lst:HypernymActivation}.

It can be seen that the function recursively activates the synsets of all hypernyms until, the model gives a result of less than zero. As the function traverses through each hypernym, the amount they are activated by decreases, favouring less general synsets.

Upon activation, the synset in question's activation is increased according to the model:

The same synset can be activated multiple times within the same sentence. In this case, a similar model to \hyperref[sec:MattBurke]{that proposed by M. Burke} \cite{MattBurkePrevious} is used: 
\[H = \sum\limits_{i=1}^N S_i\] 
The total acivation increase is equal to the sum of all synset activations within the sentence.

In Listing \ref{lst:HypernymActivation},there exists a function, hypernymModel. This function is responsible for reducing the activation increase of hypernyms as they become more general (closer to the top of the tree). This function should favour most heavily, less general hypernyms, i.e. those activated by the first few recursions of activateHypernyms.

\subsection{Disambiguation}
\label{sec:ImplementedDisambiguation}
Up to this point, we have only discussed the processes involved with changing the contents of the memory structures. The purpose of these memory structures is to provide the surrounding context of a word, to aid in its disambiguation.

@@

\subsubsection{Context}
\label{sec:DisambiguationContext}
The context required for disambiguation exists in the STS. Initially, the disambiguate function traverses through the STS, searching for a synset of the word to be disambiguated. If successful, the function will return the found synset. If unsuccessful, more investigation is required, through the use of hyponyms.

@@

This function traverses through all hyponyms of a given synset (searchItem) until either, a match with a synset in the synsetList is found, or the function reaches the base of the tree. If a match is found, the matched synset is returned as the correct word meaning. A meaning will only be found if the surrounding context of a word gives enough indication of its meaning, that is, to comply with the multiple access model, where if strong context exists, it is used for disambiguation \cite{PsychologyOfLanguage}.

@@

As stated previously, in cases where context cannot provide a viable synset, the above are used for disambiguation.

>>>>

\documentclass[]{article}
\usepackage{cite}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{tcolorbox}
\usepackage{tikz}
\usepackage{tikz-qtree}

@@

\begin{itemize}
	\item \textbf{Document}, is a list of sentences,
	\item \textbf{Sentences}, each of which is a list of words.
\end{itemize}

@@

The corpus is read a sentence at a time, for each word in the sentence a series of activations takes place. Initially, the synsets of each word are activated, followed by their hypernyms following the algorithm outlined in Listing \ref{lst:HypernymActivation}.

The function recursively traverses the tree of hypernyms an example of which is shown in Figure \ref{fig:hypernymTree}, activating each synset until, the model gives a result of less than zero. As the function traverses through each hypernym, the amount they are activated by decreases, favouring less general synsets.

\begin{figure}[h]
\begin{tikzpicture}[grow=right, sibling distance=5pt, level distance=100pt]
\Tree [.dreams [.dream.n.01 [.imagination.n.02 representational{\_}process.n.01 ] ] [.dream.n.02 [.imagination.n.01 creativity.n.01 ] ] [.ambition.n.01 [.desire.n.01 feeling.n.01 ] ] [.pipe{\_}dream.n.01 [.fantasy.n.01 imagination.n.01 ] ] [.dream.n.05 [.reverie.n.02 abstractedness.n.01 ] ] [.dream.n.06 [.perfection.n.01 state.n.02 ] ] [.dream.v.01 [.imagine.v.01 create{\_}by{\_}mental{\_}act.v.01 ] ] [.dream.v.02 [.perceive.v.01 ] ] ]
\end{tikzpicture}
\caption{HYPERNYM TREE OF "DREAMS"}
\label{fig:hypernymTree}
\end{figure}

\begin{figure}[h]
	\includegraphics[scale=0.6]{HypernymActivation.png}
	\caption{HYPERNYM ACTIVATION}
	\label{fig:HypernymAct}
\end{figure}

Upon activation, the synset in question's activation is increased by an amount, dependant upon the hypernym model used (discussed later in this section). The same synset can be activated multiple times within the same sentence, an example of which is shown in Figure \ref{fig:HypernymAct}. In this case, a similar model to \hyperref[sec:MattBurke]{that proposed by M. Burke} \cite{MattBurkePrevious} is used: 
\[H = \sum\limits_{i=1}^N S_i\] 
The total acivation increase is equal to the sum of all synset activations within the sentence.

It can be seen that, in Listing \ref{lst:HypernymActivation},there exists a function, hypernymModel. This function is responsible for reducing the activation increase of hypernyms as they become more general (closer to the top of the tree). This function should favour most heavily, less general hypernyms, i.e. those activated by the first few recursions of activateHypernyms.

@@

Upon activation, the synset in question may enter the memory structures, if it posesses a high enough activation. Both an STS and Episodic Buffer have been implemented, using python classes for each. Interactions between the two memory structures are handled by a Memory Controller. As discussed in the \hyperref[Wordnet]{WordNet section}, WordNet provides a good model of the semantic memory \cite{WN1Introduction},described in the \hyperref[LongTerm]{Long-term Store section}. For this reason it has been used heavily.

@@

The purpose of the episodic buffer, is to maintain a list of synsets which have occurred in the STS previously. This allows future activations of synsets which have already occured to receive a boost when activated in the future. Interactions with the Episodic buffer only occcur during the activation of synsets, therefore, all interactions are handled by the Memory Controller.

\subsubsection{Semantic Memory}
\label{sec:ImplementedSemanticMemory}

The Semantic Memory is responsible for maintaining all the semantic knowledge of the system. It was decided that WordNet provided the best solution, with its in depth knowledge of synsets and their relations to eachother \cite{WN2Nouns, WN4Verbs}.

NLTK's built in WordNet reader \cite{NLTK} was used to interact with WordNet. Throughout the system, synsets are used. Each of these is a reference to a synset present in WordNet, allowing their semantic relationships (accessed through their methods) to be used.

@@

When a synset is forgotten, its activation may fall below a threshold, varied in experimentation. If this occurs, the synset is removed from the STS and, as with the sendToStm function, is added to the Episodic Buffer. This prevents synsets irrelevant to the current context from existing in the STS.

\subsection{Disambiguation}
\label{sec:ImplementedDisambiguation}
Up to this point, we have only discussed the processes involved with changing the contents of the memory structures. The purpose of these memory structures is to provide the surrounding context of a word, to aid in its disambiguation. There exists a delay between the processes described previously (i.e. the initial reading of the sentence) and the disambiguation phase. This delay allows inappropriate synsets to leave the STS before they are used. The duration of this delay is to be adjusted during experimentation.

@@

\subsubsection{Context}
\label{sec:DisambiguationContext}
The context required for disambiguation exists in the STS. Two algorithms for disambiguation using this context have been implemented, one using hyponyms and one using hypernyms.

@@

This function traverses through all hyponyms of a given synset (searchItem) until either, a match with a synset in the synsetList is found, or the function reaches the base of the tree. If a match is found, the matched synset is returned as the correct word meaning.

Another algorithm, shown in Listing \ref{lst:hypernymSearch}, for disambiguation has also been implemented, using the lowest{\_}common{\_}hypernym method implemented in NLTK's WordNet reader \cite{NLTK}. This algorithm instead searches the hypernyms of both the word to disambiguate and the item in the STS, to find whether a common hypernym exists. If a common hypernym does exist, it is checked to ensure it is not too general (e.g. the common hypernym could be "entity" which is a hypernym of all nouns), and the synset is returned as the correct meaning.

\begin{lstlisting}[numbers=left, numberstyle=\small, caption={THE HYPERNYMSEARCH FUNCTION}, captionpos=b, label={lst:hypernymSearch}]
FUNCTION hypernymSearch(synsetList, searchItem):
    FOR synset in synsetList LOOP
        common_hypernym = synset.lowest_common_hypernym(searchItem)
        IF common_hypernym.depth > 4 THEN
            RETURN synset
       	END IF
   	END LOOP
\end{lstlisting}

Using either method, a meaning will only be found if the surrounding context of a word gives enough indication of its meaning, complying with the multiple access model, where if strong context exists, it is used for disambiguation \cite{PsychologyOfLanguage}.

@@

As stated previously, in cases where context cannot provide a viable synset, the above is used for disambiguation.
